---
title: "Predicting the Weather"
output:
  html_document:
    theme: cerulean
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
library(pander)
library(DT)
weather_data <- read.csv("../Data/weather.csv")
```

## Background

We were given the task of predicting the high temperature of this coming monday, the 23rd of September. We were given free range of picking our independent variable. I collected the mean humidity for for the previous week, the 16th to the 20th every year since 2011 up until 2019. 


## Analysis

$$
  \underbrace{Y_i}_\text{High Temperature} = \beta_0 + \beta_1 \underbrace{X_i}_\text{Average Humidity} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2) 
$$

This analysis is predicting the high temperature of September 23rd, 2019. Our data has collected the high temperature of the September 23rd since 2011. We have also collected the average humidity of the September 16th to the 20th going back to 2011 up until this year. We are hoping that with the collected mean humidity and corresponding high temperature of each 23rd, we can predicted what the high temperature is going to be this coming monday, the 23rd.

```{r}
weather.lm <- lm(high.temp ~ average.humidity, data = weather_data)
```

## Visualization

```{r}
plot(high.temp ~ average.humidity, xlab = "High Temperature", ylab = "Average Humidity (16th-20th)" ,data = weather_data)
abline(weather.lm)
```

## Regression

```{r}
pander(summary(weather.lm))
```

## Interpretation

```{r}
par(mfrow=c(1,3))
plot(weather.lm, which=1:2)
plot(weather.lm$residuals, main="Residuals vs Order", xlab="",
     ylab="Residuals")
```

As we look at a combination of our summary of th regression, the plot, and the residuals vs fitted graph, we can see that our regression line is not as accurate as could be.  $p = 0.9268$, this is pretty significant and a large p value. This was caused because of the lack of data we had. Since our resource only showed the past weather up until 2011 we couldn't collect that much data. That means I was only allowed to collect 8 data points. I would much rather of liked to collect 30+ data points so that my regression could be as accurate as possible. We have to step back for a second and realize that we are trying to predict the weather, thats a very tricky thing to do. Weather isn't linear which means it's not easy to predict, especially with the resources that were available. 

## Prediction

The average humidity of this week has been 43.6%. This means that if we plug that into our regression it would prompt us with a high temperature of 68.58784 degrees. So I predict that the high temperature on Monday, September 23rd, will be 69 degrees.

## Prediction Interval

```{r}
w.lm <- lm(average.humidity ~ high.temp, data = weather_data)
plot(average.humidity ~ high.temp, data = weather_data, xlim=c(0,100))
abline(w.lm)
abline(h=predict(w.lm, data.frame(high.temp = 68), interval = "prediction"), lty=2, col ="pink", v=68)
abline(h=predict(weather.lm, data.frame(average.humidity = 43.6), interval = "prediction"), lty=2, col ="pink", v=43.6)
```

According to my prediction line using the linear regression it predicts that the weather would be between 45 and 70 degrees. This is very broad because my model was very poor. It has a very low R squared value which meant that it didn't have an accurate prediction.
